---
title: "Statistics for a data rich world"
author: "Emily Faul"
date: "2023-09-12"
output: html_document
--- 


```{r}
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=3.5) 
```


```{r}
# PROGRAMMING CHALLENGE: Stats Data Rich World
library(dplyr)
```


```{r}
# Read the Dataset sheet into “R”. The dataset will be called "data".
data <- read.csv("https://covid.ourworldindata.org/data/owid-covid-data.csv",
                 na.strings = "", header=T)


```

```{r}
# Simulating the average daily temperature column for our nonsensical analysis.
# Assigning random temperature values between -10 and 35 (Celsius) to the new column.
set.seed(1234)
data$average_daily_temperature <- runif(nrow(data), -10, 35)
res <- NULL
for (i in 1:length(unique(data$date))) {
  data1 <- data[which((data$date == unique(data$date)[i]) & (data$continent == "Europe")),]
  data1 <- data1 %>% select("iso_code", "date", "new_cases", "average_daily_temperature")
  if (sum(rowSums(!is.na(data1[,3:4])) == 2) >= 10) {
    res <- rbind(res, c(unique(data$date)[i],
                        cor.test(data1[,3], data1[,4])$estimate,
                        cor.test(data1[,3], data1[,4])$p.value))
  }
}
res[which((as.numeric(res[,2]) > 0) & (as.numeric(res[,3]) <= 0.05)),]
```







```{r}
# Simulating the average daily temperature column for our nonsensical analysis.
# Assigning random temperature values between -10 and 35 (Celsius) to the new column.
set.seed(1234)
data$average_daily_temperature <- runif(nrow(data), -10, 35)
res <- NULL
for (i in 1:length(unique(data$date))) {
  data1 <- data[which((data$date == unique(data$date)[i]) & (data$continent == "Europe")),]
  data1 <- data1 %>% select("iso_code", "date", "new_cases", "average_daily_temperature")
  if (sum(rowSums(!is.na(data1[,3:4])) == 2) >= 10) {
    res <- rbind(res, c(unique(data$date)[i],
                        cor.test(data1[,3], data1[,4])$estimate,
                        cor.test(data1[,3], data1[,4])$p.value))
  }
}
res[which((as.numeric(res[,2]) > 0) & (as.numeric(res[,3]) <= 0.05)),] 
#1st column represents: A unique date.
#2nd column represents: The estimated correlation coefficient between the two variables (new cases and average daily temperature).
#3rd column represents: The p-value associated with the correlation test.
```


To demonstrate how data can be "cherry picked" we generated a uniform distribution and generated a list of values between -10 and 35˚C to simulate average daily temperature data. Following statistical analysis we found on multiple dates (listed below) average daily temperature values are positively correlated to new COVID cases in Europe.  

*First ten dates generated show significant correlation. 

 [1,] "2020-03-27" "0.283349810738522" "0.0439195033699029" 
 [2,] "2020-09-09" "0.283560998528383" "0.043755574450734"  
 [3,] "2020-10-24" "0.281729177478701" "0.0451942674483055" 
 [4,] "2020-11-28" "0.310802610790601" "0.0264293749046018" 
 [5,] "2021-04-01" "0.314722691341704" "0.0244904617628498" 
 [6,] "2021-04-15" "0.29398371143264"  "0.0362659730622546" 
 [7,] "2021-08-21" "0.277315088031726" "0.048820224619938"  
 [8,] "2021-10-02" "0.322815724561232" "0.0208632792934312" 
 [9,] "2021-12-05" "0.278494210356606" "0.0478292321606055" 
[10,] "2022-02-15" "0.286350952123346" "0.0416365542255234" 









  




